---
title: "Course_project1"
author: "David Romero"
date: "1/15/2021"
output: html_document
---

## Code for reading in the dataset and/or processing the data
```{r, echo=TRUE}
dir()
unzip("repdata_data_activity.zip")
library(dplyr)
library(lubridate)
activity_data <- read.csv("activity.csv")
head(activity_data)
activity_data$date <- ymd(activity_data$date)
```


## Histogram of the total number of steps taken each day
```{r,echo=TRUE}
steps_per_day <- aggregate(steps ~ date, activity_data, sum, na.rm = TRUE)
hist(steps_per_day$steps, xlab = "Steps", ylab = "Number of Days", main = "TOTAL OF STEPS")
```

## Mean and median number of steps taken each day
```{r, echo=TRUE}
mean(steps_per_day$steps, na.rm = TRUE)
median(steps_per_day$steps, na.rm = TRUE)
```

## Time series plot of the average number of steps taken
```{r, echo=TRUE}
averagesteps <- aggregate(steps ~ interval, activity_data, mean)
with(averagesteps, plot(interval, steps, type = "l"))
```

## The 5-minute interval that, on average, contains the maximum number of steps
```{r, echo=TRUE}
averagesteps[which.max(averagesteps[,2]), 1]
```

## Code to describe and show a strategy for imputing missing data
1. Calculate and report the total number of missing values in the dataset 
```{r, echo=TRUE}
missing_data <- is.na(activity_data[,1])
```

2. Devise a strategy for filling in all of the missing values in the dataset (mean)
```{r, echo=TRUE}
meansteps <- mean(averagesteps$steps)
```

3. Create a new dataset that is equal to the original dataset bur with the missing data filled in
```{r, echo=TRUE}
activity_full <- activity_data
activity_full[missing_data, 1] <- meansteps
head(activity_full)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r, echo=TRUE}
full_steps_per_day <- aggregate(steps ~ date, activity_full, sum)
hist(full_steps_per_day$steps, xlab = "Steps", ylab = "Number of Days", main = "TOTAL OF STEPS")
mean(full_steps_per_day$steps, na.rm = TRUE)
median(full_steps_per_day$steps, na.rm = TRUE)
```
The mean and didn´t change after the replacements of NA`s, the median changed about 0,1%.

## Panel plot comparing the average number of steps taken per 5-minute inferval acrros weekdays and weekends
```{r, echo=TRUE}
library(ggplot2)
activity_full$weekday <- weekdays(activity_full$date)
activity_full$DateType <- ifelse(activity_full$weekday == "Saturday" | activity_full$weekday == "Sunday", "Weekend", "Weekday")
steps_per_time <- aggregate(steps ~ interval + DateType, data = activity_full, mean)
ggplot(steps_per_time, aes(interval, steps)) + geom_line(col= "blue") + ggtitle("Average steps weekdays vs. weekends") + xlab("Time") + ylab("Steps") + theme(plot.title = element_text(face= "bold", size = 12)) + facet_grid(DateType ~.)
```

